---
title: "Project1 - Group5"
author: "Group-5 Anshita Aishwarya, Shrutika Swamy"
date: "11/24/2021"
output: pdf_document
---

## Importing the required libraries

```{r Importing Libraries}
library(dplyr)
library(factoextra)
library(cluster)
library(clValid)
library(ggplot2)
library(ggpubr)
# library(plot3D)
library(plotly)
```


## World Indicators - Data Prep and Cleaning

```{r Data Prep and Cleaning}

# Reading of data 
world_indicator <- read.csv("World Indicators.csv")

# Labeling the different rows with the country names
rownames(world_indicator) <- world_indicator$Country

# Removing dollar sign, percent symbols and commas
world_indicator$GDP = as.numeric(gsub("[\\$,]", "", world_indicator$GDP))
world_indicator$Health.Exp.Capita = as.numeric(gsub("[\\$,]", "", world_indicator$Health.Exp.Capita))
world_indicator$Business.Tax.Rate = as.numeric(gsub("[\\%,]", "", world_indicator$Business.Tax.Rate))

# Dropping the non numerical columns
newdf <- world_indicator[, 1:18]
df_actual <- world_indicator

# Dropping the columns that have more than 70 NA values
newdf <- newdf[ , colSums(is.na(newdf)) < 70]

# Removing the non-significant columns
newdf <- newdf[, -c(2,3,7,9,12,13)]

# Omitting the rows with NA values
newdf <- na.omit(newdf)

# Scaling of data
newdf <- as.data.frame(scale(newdf))

```

## Performing K-Means Clustering

```{r K-Means Clustering}
# This clustering is obtained by considering all the attributes including birth rate, GDP, health expenditure, infant mortality rate, life expectancy (males and females) and population (different age groups)

# Elbow method: Determining optimal number of clusters
fviz_nbclust(newdf, kmeans, method = "wss")
fviz_nbclust(newdf, kmeans, method = "silhouette")

# All the methods above report k = 2 as the optimal number of clusters

# Using k-means clustering with k=2
km_world_indi <- kmeans(newdf, 2, nstart = 20)
km_world_indi$cluster
summary(km_world_indi)

# K-means clustering of World Indicator data
clusplot(newdf, km_world_indi$cluster, main = '2D representation of the Cluster solution',
         color = TRUE, shade = TRUE, labels = 1, lines = 1)

```

## Performing Hierarchical Clustering

```{r Hierarchical Clustering}

# This clustering is obtained by considering all the attributes including birth rate, GDP, health expenditure, infant mortality rate, life expectancy (males and females) and population (different age groups)

hc_world_indi <- hclust(dist (newdf), method = "average")
summary(hc_world_indi)
plot(hc_world_indi,main = "Clustered Dendogram of World Indicator Data")

# Using k = 3
hcm_world_indi <- cutree(hc_world_indi, k = 3)
plot(hcm_world_indi, main = "Solution by Hierarchical Clustering of World Indicator")

```

## Performing Internal Validation Metrics to evaluate the performance of clustering

1.) Connectivity - For a particular clustering partition C = {C1,...,CK} of the N observations into K disjoint clusters, the connectivity is defined as summation of x(i) (i ranging from 1 to N, total number of observations) multiplied with summation of nn(i,j)
(for i ranging from 1 to N and j ranging from 1 to L, parameter giving the number of nearest neighbors to use. Based on the different internal validation measures, the  connectivity has a value between zero and ∞ and should be minimized.

2.) Silhouette Width - The Silhouette Width is the average of each observation’s Silhouette value. For observation i, it is defined as the ratio of (bi − ai) and maximum value of (bi, ai), where ai is the average distance between i and all other observations in the same cluster, and bi is the average distance between i and the observations in the nearest neighboring cluster.Silhouette Width lies in the interval [−1, 1] and should be maximized.

3.) Dunn Index - The Dunn Index is the ratio of the smallest distance between observations not in the same cluster to the largest intra-cluster distance.Dunn Index has a value between zero and ∞ and should be maximized for determining better cluster quality.

```{r Internal Validation}
# Since our data does not have labels, we cannot perform external validation

int_validate <- clValid(newdf, nClust = 2:6, clMethods = c("kmeans", "hierarchical"),
                    validation = "internal")
summary(int_validate)

print(paste0("Thus it is evident from these calculated validation measures that hierarchical clustering has a better cluster quality as compared to k-means"))

```

## Comparing clustering solutions and grouping countries based on clustering methods

```{r Comparing clustering solutions}

# Comparing clustering solutions

# It has been depicted in the previous question based on the calculated validation measures that hierarchical clustering has a better cluster quality as compared to k-means"))

# Grouping the list of the countries clustered via k-means
listofcluster_kmeans = order(km_world_indi$cluster)
grouped_countries_kmeans <- data.frame(kmeans_cluster = km_world_indi$cluster[listofcluster_kmeans])
View(grouped_countries_kmeans)

# Grouping the list of the countries clustered via hierarchical
listofcluster_hier = order(hcm_world_indi)
grouped_countries_hier <- data.frame(hierarchical_cluster = hcm_world_indi[listofcluster_hier])
View(grouped_countries_hier)

```

## Data Plotting

```{r Data Plotting}

# This clustering is obtained considering 2 attributes - GDP and infant mortality rate
# Scatter Plot 1: Infant Mortality Rate v/s GDP (using k-means clustering)
newdf_scatter1 <- world_indicator[, c(5, 9)]
newdf_scatter1 <- na.omit(newdf_scatter1)
newdf_scatter1_scaled <- as.data.frame(scale(newdf_scatter1))

fviz_nbclust(newdf_scatter1_scaled, kmeans, method = "silhouette")
km_scatter1 <- kmeans(newdf_scatter1_scaled, 2, nstart = 20)
km_scatter1$cluster
summary(km_scatter1)

plot(newdf_scatter1$Infant.Mortality.Rate, newdf_scatter1$GDP, 
     main = "Infant Mortality Rate v/s GDP",
     xlab = "Infant Mortality Rate",
     ylab = "GDP", 
     col = km_scatter1$cluster, pch = 19, frame = FALSE)

# This clustering is obtained considering 2 attributes - Urban population and energy usage
# Scatter Plot 2: Urban Population v/s Energy Usage (using k-means clustering)
newdf_scatter2 <- world_indicator[,c(4,18)]
newdf_scatter2 <- na.omit(newdf_scatter2)
newdf_scatter2_scaled <- as.data.frame(scale(newdf_scatter2))

fviz_nbclust(newdf_scatter2_scaled, kmeans, method = "silhouette")
km_scatter2 <- kmeans(newdf_scatter2_scaled, 10, nstart = 20)
km_scatter2$cluster
summary(km_scatter2)

plot_ly(newdf_scatter2,
        x = ~Population.Urban,
        y = ~Energy.Usage,
        color = ~km_scatter2$cluster) %>% 
  layout(title = "Urban Population v/s Energy Usage",
         xaxis = list(title = "Urban Pouplation"),
         yaxis = list(title = "Energy Usage"))

# This clustering is obtained considering 2 attributes - Population 15-64 years and internet usage
# Scatter Plot 3: Population 15-64 years v/s Internet Usage (using hierarchical clustering)
newdf_scatter3 <- world_indicator[, c(10, 16)]
newdf_scatter3 <- na.omit(newdf_scatter3)
newdf_scatter3_scaled <- as.data.frame(scale(newdf_scatter3))

hc_scatter3 <- hclust(dist (newdf_scatter3_scaled), method = "average")
summary(hc_scatter3)
plot(hc_scatter3)

#Using k=4
hcm_scatter3 <- cutree(hc_scatter3, k = 4)
plot(hcm_scatter3)
newdf_scatter3$Class_hier <- hcm_scatter3

ggplot(newdf_scatter3, aes(x = Population.15.64, y = Internet.Usage, color = as.factor(Class_hier))) +
  geom_point(stat = "identity") +
  ggtitle("Population 15-64 years v/s Internet Usage") +
  xlab("Pouplation 15-64 years") +
  ylab("Internet Usage") +
  labs(color = "Cluster")+
  scale_colour_brewer(palette = "Set1") +
  theme(plot.title = element_text(size = 13, face = "bold.italic", color = "black", hjust = 0.5),
        axis.title.x = element_text(size = 7, face = "bold"),
        axis.title.y = element_text(size = 7, face = "bold"),
        axis.text.x = element_text(size = 5, angle = 0),
        axis.text.y = element_text(size = 5, angle = 90),
        legend.position = "right",
        legend.title = element_text(size = 7, face = "bold"),
        legend.text = element_text(size = 7, face = "bold"),
        strip.text = element_text(size = 7))
```